Polynomial phase estimation by least squares phase unwrapping

### 2013-12-16 paper.tex
added plots and modified simulation section to include product CPF-HAF

This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.  The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.  In all simulations the unknown amplitude is $\rho = 1$ and $N = 199$.  The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.  The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.  The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\tilde{\mubf} - \widehat{\mubf}_t)$ are computed.  The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.  In all simulations four \emph{lag sets} are used for the PHAF estimator~\cite{Barbarossa_PHAF_1998} and the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.  The maximum number of branches allowed is $K=20N$.  The arithmetic complexity of the LSU estimator is $O(N^3\log N)$ in this case.

@@

Figure~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.  In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.  In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF, PHAF and CPF-HAF estimators.  As in the case where $m=3$, the region of coefficients suitable for the HAF, PHAF and CPF-HAF estimators is much smaller that the identifiable region.  The effect becomes more severe as $m$ increases.  When $m=5$ and $N=199$ the space of coefficients searched by these estimators is approximately $10^{-16}$ the size of the identifiable region.  This is a dramatic reduction in the range of detectable polynomial phase signals. Figure~\ref{plot:m5big} displays results under the same conditions as Figure~\ref{plot:m5} but with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.  The conclusion is similar to that from Figure~\ref{plot:m3big}.  The performance of the LSU estimator remains unchanged, but the HAF, PHAF and CPF-HAF estimators fail completely.  The Zhou and Wang estimator can again be used to increase the range of parameters for the HAF, but simulations indicate that the threshold for the estimator is approximately~\unit[103]{dB}.

Tables~\ref{tablm1},~\ref{tablm3} and~\ref{tablm5} show the values of SNR at which the estimators reach a threshold for polynomial phase signals of order $m=1,3$ and $5$.  The values in the tables are approximated using Monte-Carlo simulation with $X_1, \dots, X_N$ pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian.  When $m=3$ the true coefficients are as in Figure~\ref{plot:m3small} and when $m=5$ the true coefficients are as in Figure~\ref{plot:m5} so that the coefficients are inside the range acceptable for the HAF, CPF and CPF-HAF estimators.  When $m=1$ (single frequency estimation) the LSU estimator can be computed exactly using the algorithm from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  In this case the value at which the LSU estimator reaches a threshold is similar to the maximum likelihood periodogram estimator~\cite{Quinn2001,Hannan1973,Rife1974}.  This property was observed in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  Both the LSU and periodogram estimators are strongly consistent, and so, the threshold values can be made arbitrarily small by choosing $N$ large.  This is indicated in the bottom row of Table~\ref{tablm1} by the $-\infty$ symbols.  When $m>1$ the LSU estimator can be exactly computed using the sphere decoder~\cite{Agrell2002,Viterbo_sphere_decoder_1999} in a reasonable amount of time when $N < 60$.  When $N \geq 60$ we use the approximate $K$-best algorithm.  We observe that the $K$-best algorithm produces estimates with the same threshold as the exactly computed LSU estimator when $N = 19$ and $49$.  As $N\rightarrow\infty$ the asymptotic threshold of the HAF estimator has been studied in~\cite{Porat_asympt_HAF_DPT_1996} and these values are indicated in the bottom rows of Tables~\ref{tablm3} and~\ref{tablm5}.  By Theorem~\ref{thm:asymp_proof} the LSU estimator is strongly consistent and the threshold can be made as small as desired by choosing $N$ large.  Whether the LSU estimator computed approximately using the $K$-best algorithm is strongly consistent is an open theoretical question.  That the $K$-best algorithm provides a close approximation of the exact LSU estimator for small $N$ is an encouraging observation.

@@

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
$N$ & LSU & LSU ($K$-best) & CPF & HAF & ZW \\ \toprule
$19$ & 10 & 10 & 8 & 11 & 31 \\ 
$49$ & 7 & 7 & 5 & 9 & 29 \\ 
$99$ & & 6 & 3 & 7 & 29 \\ 
$199$ & & 3 & 1 & 5 & 28 \\ 
$299$ & & 1 & -1 & 5 & 27 \\ 
$\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=3$}
\label{tablm3}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
$N$ & LSU & LSU ($K$-best) & CPF-HAF & HAF & ZW \\ \toprule
$19$ & 22  & 22 & 17 & 23 & $>100$ \\ 
$49$ & 17 & 17 & 14 & 21 & $>100$  \\ 
$99$ &  & 15 &  12 &  19 & $>100$ \\ 
$199$ & & 12 & 10 & 18 & $>100$ \\ 
$299$ &  & 11 & 9 & 17 & $>100$ \\ 
$\infty$ & $-\infty$ &  &  & 16.6 &  \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=5$}
\label{tablm5}
\end{table}

>>>>

This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the product high order ambiguity function (PHAF) estimator~\cite{Barbarossa_PHAF_1998}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimators that result from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.  The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.  In all simulations the unknown amplitude is $\rho = 1$ and $N = 199$.  The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.  The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.  The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\tilde{\mubf} - \widehat{\mubf}_t)$ are computed.  The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.  In all simulations four \emph{lag sets} are used for the PHAF estimator~\cite{Barbarossa_PHAF_1998} and the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.  The maximum number of branches allowed is $K=20N$.  The arithmetic complexity of the LSU estimator is $O(N^3\log N)$ in this case.

@@

Figure~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.  In this case the LSU and PHAF estimators reach a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, the estimator that combines the CPF and HAF at \unit[10]{dB}, and the product form of this estimator (PCH) at \unit{6}[dB] (see~\cite[Section~5.A]{Djurovic_haf_cpf_2012}).  In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF, PHAF, CPF-HAF and PCH estimators.  As in the case where $m=3$, the region of coefficients suitable for the HAF, PHAF, CPF-HAF and PCH estimators is much smaller than the identifiable region.  The effect becomes more severe as $m$ increases.  When $m=5$ and $N=199$ the space of coefficients searched by these estimators is approximately $10^{-16}$ the size of the identifiable region.  This is a dramatic reduction in the range of detectable polynomial phase signals. Figure~\ref{plot:m5big} displays results under the same conditions as Figure~\ref{plot:m5} but with $\tilde{\mubf} = \left[\tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.  The conclusion is similar to that from Figure~\ref{plot:m3big}.  The performance of the LSU estimator remains unchanged, but the HAF, PHAF, CPF-HAF and PCH estimators fail completely.  The Zhou and Wang estimator can again be used to increase the range of parameters for the HAF, but simulations indicate that the threshold for the estimator is approximately~\unit[103]{dB}.

Tables~\ref{tablm1},~\ref{tablm3} and~\ref{tablm5} show the values of SNR at which the estimators reach a threshold for polynomial phase signals of order $m=1,3$ and $5$.  The values in the tables are approximated using Monte-Carlo simulation with $X_1, \dots, X_N$ pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian.  When $m=3$ the true coefficients are as in Figure~\ref{plot:m3small} and when $m=5$ the true coefficients are as in Figure~\ref{plot:m5} so that the coefficients are inside the range acceptable for the HAF, CPF, CPF-HAF and PCH estimators.  When $m=1$ (single frequency estimation) the LSU estimator can be computed exactly using the algorithm from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  In this case the value at which the LSU estimator reaches a threshold is similar to the maximum likelihood periodogram estimator~\cite{Quinn2001,Hannan1973,Rife1974}.  This property was observed in~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}.  Both the LSU and periodogram estimators are strongly consistent, and so, the threshold values can be made arbitrarily small by choosing $N$ large.  This is indicated in the bottom row of Table~\ref{tablm1} by the $-\infty$ symbols.  When $m>1$ the LSU estimator can be exactly computed using the sphere decoder~\cite{Agrell2002,Viterbo_sphere_decoder_1999} in a reasonable amount of time when $N < 60$.  When $N \geq 60$ we use the approximate $K$-best algorithm.  We observe that the $K$-best algorithm produces estimates with the same threshold as the exactly computed LSU estimator when $N = 19$ and $49$.  As $N\rightarrow\infty$ the asymptotic threshold of the HAF estimator has been studied in~\cite{Porat_asympt_HAF_DPT_1996} and these values are indicated in the bottom rows of Tables~\ref{tablm3} and~\ref{tablm5}.  By Theorem~\ref{thm:asymp_proof} the LSU estimator is strongly consistent and the threshold can be made as small as desired by choosing $N$ large.  Whether the LSU estimator computed approximately using the $K$-best algorithm is strongly consistent is an open theoretical question.  That the $K$-best algorithm provides a close approximation of the exact LSU estimator for small $N$ is an encouraging observation.

@@

\begin{table}[t]
\centering
\begin{tabular}{lrrrrr}
$N$ & LSU & $K$-best & CPF & HAF & ZW \\ \toprule
$19$ & 10 & 10 & 8 & 11 & 31 \\ 
$49$ & 7 & 7 & 5 & 9 & 29 \\ 
$99$ & & 6 & 3 & 7 & 29 \\ 
$199$ & & 3 & 1 & 5 & 28 \\ 
$299$ & & 1 & -1 & 5 & 27 \\ 
$\infty$ & $-\infty$ & & & 3.77 & \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=3$}
\label{tablm3}
\end{table}

\begin{table}[t]
\centering
\begin{tabular}{lrrrrrr}
$N$ & LSU & $K$-best & CPF-HAF & PCH  & HAF & ZW \\ \toprule
$19$ & 22  & 22 & 17 & 13 & 23 & $>100$ \\ 
$49$ & 17 & 17 & 14 & 10 & 21 & $>100$  \\ 
$99$ &  & 15 &  12 &  8 & 19 & $>100$ \\ 
$199$ & & 12 & 10 & 6 & 18 & $>100$ \\ 
$299$ &  & 11 & 9 & 5 & 17 & $>100$ \\ 
$\infty$ & $-\infty$ &  &  & & 16.6 &  \\ \bottomrule
\end{tabular}
\caption{Estimated SNR threshold in decibels with $m=5$}
\label{tablm5}
\end{table}
